---
title: "assignment4packagesdevgur1"
format: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# assignment4packagesdevgur1

<!-- badges: start -->
<!-- badges: end -->

The goal of this package is to provide an interactive way to explore and communicate insights from the Australian Bushfire dataset used in earlier assignments for ETC5523.
It combines data, analysis functions, and a Shiny dashboard to help users visualise bushfire trends and understand key relationships between fire weather, temperature anomalies, and geographic regions.

## Features

- Contains the dataset a2_bushfire as an .rda object

- Includes preprocessing and cleaning scripts inside data-raw/

- Provides a Shiny app (inst/app/app.R) for interactive visualisation

- Includes documentation generated with roxygen2

- Fully installable as an R package and deployable with pkgdown


## Installation

You can install the development version of assignment4packagesdevgur1 from [GitHub](https://github.com/) with:

```{r}
# install.packages("pak")
pak::pak("ETC5523-2025/assignment-4-packages-and-shiny-apps-devgur1")
```

After installation, you can load the package and launch the app:

```{r}
library(assignment4packagesdevgur1)
run_app()

```
This will open the Shiny app in your browser.

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(assignment4packagesdevgur1)

# Load the bushfire dataset
data("a2_bushfire")

# Display summary
summary(a2_bushfire)
```

## The Shiny App

The interactive dashboard allows you to:

- Select regions or years to explore bushfire data

- Visualise key trends in temperature, rainfall, and fire severity

- Interpret results through descriptive text panels

To run it directly:
```{r}
assignment4packagesdevgur1::run_app()

```

## pkgdown Site
A complete documentation website (functions, data, vignette, and app details) is available at:
https://ETC5523-2025.github.io/assignment-4-packages-and-shiny-apps-devgur1

## License
This package is licensed under the MIT License.

## Author
Dev Karan Gur
Monash University – ETC5523: Communicating with Data
Semester 2, 2025

