---
title: "Exploring Australian Bushfire Data with R"
format: gfm
--- 

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
<!-- badges: end -->

The goal of this package is to provide an interactive way to explore and communicate insights from the Australian Bushfire dataset used in earlier assignments for ETC5523-Communicating with Data.
It combines data, analysis functions, and a Shiny dashboard to help users visualise bushfire trends and understand key relationships between fire weather, temperature anomalies, and geographic regions.

## Features

- Contains the dataset a2_bushfire as an .rda object

- Includes preprocessing and cleaning scripts inside data-raw/

- Provides a Shiny app (inst/app/app.R) for interactive visualisation

- Includes documentation generated with roxygen2

- Fully installable as an R package and deployable with pkgdown


## Installation

You can install the development version of assignment4packagesdevgur1 from [GitHub](https://github.com/) with:

```{r install, echo=TRUE}
# install.packages("pak")
pak::pak("ETC5523-2025/assignment-4-packages-and-shiny-apps-devgur1")
```

After installation, you can load the package and launch the app:

```{r, eval=FALSE, message=FALSE, warning=FALSE, echo=TRUE}
library(bushfiredataexplorer)
run_app()
```
This will open the Shiny app in the browser.

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(bushfiredataexplorer)

# Load the bushfire dataset
data("a2_bushfire")

# Display summary
summary(a2_bushfire)
``` 

## The Shiny App

The interactive dashboard allows you to:

- Select regions or years to explore bushfire data

- Visualise key trends in temperature, rainfall, and fire severity

- Interpret results through descriptive text panels

To run it directly:

```{r, eval=FALSE, echo=TRUE}
bushfiredataexplorer::run_app()

```

## Pkgdown Site
A complete documentation website (functions, data, vignette, and app details) is available at:
https://ETC5523-2025.github.io/assignment-4-packages-and-shiny-apps-devgur1
  
## License
This package is licensed under the MIT License.

## Author

Dev Karan Gur

Monash University – ETC5523: Communicating with Data


