---
title: "Getting Started with the Bushfire Data Explorer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with the Bushfire Data Explorer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  message = FALSE,
  warning = FALSE
)
```


## Introduction

The Bushfire Data Explorer is an R package designed to help users explore and communicate insights from historical Australian bushfire data.
It combines datasets (a1_media, a2_bushfire) with analysis functions and an interactive Shiny dashboard to visualise long-term fire trends, temperature anomalies, and rainfall changes from 1900 to 2020.

This vignette demonstrates how to:

1) Load the included datasets

2) Summarise and visualise key variables

3) Launch the interactive Shiny app

## Load the Package
```{r setup}
library(bushfiredataexplorer)
```

## Explore the Data
The package includes two datasets:

- a1_media – communication styles across articles and media types

- a2_bushfire – historical bushfire metrics (Fire Weather Index, Max Temperature, Rainfall)

```{r}
# Load and preview the bushfire dataset

data("a2_bushfire")
head(a2_bushfire)

# Display summary statistics

summary(a2_bushfire)

```

## Visualise Bushfire Trends

We can use ggplot2 to explore long-term changes in the Fire Weather Index (FWI):
```{r}
library(ggplot2)

ggplot(a2_bushfire, aes(x = year, y = FWI)) +
  geom_line(color = "#0072B2", linewidth = 1) +
  geom_smooth(method = "loess", se = FALSE, color = "orange") +
  theme_minimal(base_size = 13) +
  labs(
    title = "Fire Weather Index Trends in Australia (1900–2020)",
    subtitle = "Higher values indicate increased bushfire risk",
    x = "Year",
    y = "Fire Weather Index (FWI)"
  )

```

## Launch the Interactive Dashboard

```{r}
# Launch the Shiny app in your RStudio or browser

# run_app()

```

When we run the app, you can:

- Select metrics (FWI, TX7x, Rainfall)

- Adjust year range using the slider

- View dynamic visualisations and summary descriptions

## Conclusion

The Bushfire Data Explorer provides a powerful and easy-to-use framework for exploring environmental and climatic patterns linked to Australian bushfires.
It demonstrates how R packages can combine data, visualisation, and interactivity to support reproducible communication of complex data.


